---
title: "Attempt to run the ETIS code"
author: "Carl James Schwarz"
date: '`r format(Sys.time(), "%Y-%m-%d")`'
output:
  html_document:
    number_sections: yes
    toc: yes
  pdf_document:
    keep_tex: false
    number_sections: yes
    toc: yes
    fig_caption: true
    extra_dependencies: ["float"]
---

# Preliminaries.

Before running the *R* scripts for the ETIS analysis, you will need to set up
an SQL server, populate a database to be accessed by the R scripts, install JAGS, etc.

## Setup the ETIS database on your local machine.

### Windoze local computer

**To be provided later**

### Macintosh local computer

You must install a SQL server on your local computer and populate it with the ETIS records.
This document uses the *Postgres.app* SQL server, but you could use any other server. Your install
and population instructions will be similar.

#### Download the Postgres.app
 Download Postgres.app from https://www.postgresql.org/download/macosx/
Install it on your local computer in the usual way. The installation will 
create an icon (an elephant) in the mac menu bar on the top right

#### Create a new database and populate it from the *.sql file of ETIS records.

The `Postgres.app` will automatically create one or more sample databases (that are empty) when installed.

##### Launch the application and start the *psql* application

Click on any of the databases that are established to open the "psql" program in a terminal window.
I clicked on the database called ``cschwarz` (your existing database may be different)
The Terminal window should then read something like:

```
/Applications/Postgres.app/Contents/Versions/13/bin/psql -p5432 "cschwarz"
cschwarz@schwarz-iMac-2020 ~ % /Applications/Postgres.app/Contents/Versions/13/bin/psql -p5432 "cschwarz"
psql (13.1)
Type "help" for help.
cschwarz=# 
```
Notice the `cschwarz=#` prompt indicated you are in the *psql* program.

##### Create a NEW database to hold the ETIS records. 

I will now create a new database (called `etis`) to hold the ETIS database using the `create` command

```
cschwarz=#   create database etis;
CREATE DATABASE
```


##### Sign into the "etis" database.

We now sign into the `etis` database. The "\" in the command below is important

```
cschwarz=# \c etis

You are now connected to database "etis" as user "cschwarz".
```




##### Populate the new "etis" database 
We will now populate the `etis` database by importing the contents of the the *.sql file. 
See https://stackoverflow.com/questions/53245701/import-sql-file-into-postgres-using-psql for additional details.

Notice that the prompt below will have changed from `cschwarz=#` to `etis=#`. 
The path name to the *sql file will differ from the path below. 
I would drag and drop the file into the terminal window to get the full path name rather than typing it from scratch.

```
etis=# \i /Users/cschwarz/Dropbox/etis_production_06_06_2018.sql 
```


You will now get many lines of output as the `etis` database is populated from the *.sql file.
I've shown the first few and last few output lines below:

```
SET
SET
SET
...
CREATE EXTENSION
...
CREATE TABLE
```


##### Exit the database and the `psql` program

Exit from the database and `psql` program

```
etis=# \q
```


#### Accessing the your `etis` database from *R*

Be sure that the `Postgres.app` is running (click on the elephant in the menu bar) 
and that it is "started" (there is a button when you open the application).

Within your R script, use

```
host.name <- 'localhost'
pg.port <- 5432
user.name <- 'postgres'
db.name <- 'etis'
```

before connecting to the database. Happy SQLing!

## JAGS install

JAGS is a program that performs the Bayesian Analysis. It is called from *R*
but is not an *R* package and must be installed separately.

**Provide instructions here on how to do this **


# Producing the CoP18 ETIS results

```{r echo=TRUE, message=TRUE, warning=TRUE}
# Problems encountered: Code runs but need modification for your local machine
source(file.path("R","PG settings.R"))

# I had to change the code to:
host.name <- 'localhost'
pg.port <- 5432
user.name <- 'postgres'
# 
# If a password is required, enter it between the quote marks here:
passwd <- 'XXXX'
#
# ... then set the database name:
db.name <- 'etis'

```

Create a directory *ProcessedData*, *JAGS Summaries* and *2018 JAGS MCMCoutputs* to store intermediate results
and modify the paths below to point to these directories.

```{r echo=TRUE, message=TRUE, warning=TRUE}
# Problems encountered:
# (a). This code didn't run directly because it sets a working directory using static paths.
# this likely needs to be changed. I would move the setting of paths into a separate script and
# source that script so it done once for all scripts rather than modifying the code in every script.
# (b) Code needs to be repaired to use file.path(directory,file) to make it system independent etc.
# (c) people will miss that code needs to be run twice. Be explicit
# (d) Remov year.from and year.to from the files.

base.dir <- '/Users/cschwarz/Dropbox/00-current-tasks/2020-12-Mike-Elephants/ETIS_CITESReporting_RCode-v.CoP18_CS'

path.Rcode    <- file.path(base.dir, "R"             )   
path.data     <- file.path(base.dir, "ProcessedData" )
path.wkg      <- file.path(base.dir, "ProcessedData" )
path.out.sum  <- file.path(base.dir, "JAGS summaries")
path.out.mcmc <- file.path(base.dir, "2018 JAGS MCMCoutputs")

# 2. Need to launch the SQL server in the back ground.

# need to run the following file twice
year.from <- 1900
year.to   <- 2100
source(file.path(path.Rcode,"Sz data setup.R"))

year.from <- 2007
year.to   <- 2017
source(file.path(path.Rcode,"Sz data setup.R"))
```

```{r echo=TRUE, message=TRUE, warning=TRUE}
# Problems encountered:

source(file.path(path.Rcode,"Sz id weights changed.R"))
```

```{r echo=TRUE, message=TRUE, warning=TRUE}
# Problems encountered:
# (0) You also need to run the model.development first because it creates Rdata files.
# (a) this needs to be run before "wgt est.R" because  "wgt est.R" loads "wt est.Rdata" which is created by this code
# (b) Comment out the R and ProcessData paths. The path names are different here? Why?
# (c) Wants the file "'ivory sz recs 1900_2100.csv'" and not clear where this is created?
# (d) Change path.wkg to path.data or set it equal.
# (e) Need to add _tidy to csv file names created in previous program
# (f) Set the par(ask=FALSE) so code does not prompt to show the plots

par(ask=FALSE)
source(file.path(path.Rcode,"wgt est models development.R"))
```



```{r echo=TRUE, message=TRUE, warning=TRUE}
# Problems encountered:
# (a) this needs to be run before "wgt est.R" because  "wgt est.R" loads "wt est.Rdata" which is created by this code
# (b) Comment out the R and ProcessData paths. The path names are different here? Why?
# (c) Wants the file "'ivory sz recs 1900_2100.csv'" and not clear where this is created?
# (d) Change path.wkg to path.data or set it equal.
# (e) Need to add _tidy to csv file names created in previous program
source(file.path(path.Rcode,"wgt est models.R"))
```


```{r echo=TRUE, message=TRUE, warning=TRUE}
# Problems encountered:
# (a) Comment out the R and Processed data paths. see previous note.
# (b) Similarly, the year.to and year.from should be commented out and again set in another source file once for all code scripts
# (c) Need to adjust the file path for reading in the created ivory data
# (d) avoid T/F for TRUE/FALSE
# (e) wanted load('wt est.Rdata') but this doesn't exist yet. Created by "wgt est models.R" which needs to be run first.

source(file.path(path.Rcode,"Wgt est.R"))
```

```{r echo=TRUE, message=TRUE, warning=TRUE}
# Problems encountered:
# missing from GitHub sites?

#source(file.path(path.Rcode,"Wgt est models comparison.R"))   

```
```{r echo=TRUE, message=TRUE, warning=TRUE}
# Problems encountered:
# (a) changed file name to sz.. from Sz
# (b) commented out year.from and year.to as set previously.
# (c) file paths set outside code block
# (d) Changes year.to and year.from to 2008 to 2017. Next code block calls this year.crit.to year.crit.from

source(file.path(path.Rcode,"sz_inout_tables_df_gps_transit.R"))   

```



```{r echo=TRUE, message=TRUE, warning=TRUE}
# Problems encountered:
# (a) changed file name to sz.. from Sz
# (b) commented out year.from and year.to as set previously
#     ditto for year.crit.to and year.crit.from.
# (c) file paths set outside code block
# (d) csv files have _tidy suffix that needs to be added.
# (e) dplyr:: added to select
# (f) mutate problem as seen by earlier

source(file.path(path.Rcode,"sz datasets_ALL gps.R"))   

```



```{r echo=TRUE, message=TRUE, warning=TRUE}
#  R script that shows interpolation/assumptions of ETIS subsidiary database to calculate
#  covariates (missing)  3

```


```{r echo=TRUE, message=TRUE, warning=TRUE}
# Problems encountered:
# (a) year.to and year.from commented out since set outside code
# (a) path name commented out since outside code
# (b) update path name to country names

source(file.path(path.Rcode,"Covar setup.R"))   

```

```{r echo=TRUE, message=TRUE, warning=TRUE}
# Problems encountered:
# (a) year.to and year.from commented out since set outside code
# (a) path name commented out since outside code
# (b) update path name to country names

#source(file.path(path.Rcode,"Trade route calculation.R"))   
# there is still a problem with mutate that needs to be resolved.

```
```{r echo=TRUE, message=TRUE, warning=TRUE}
#14.  Checking seizure records (This file can’t be run because it compares records used in
#2017 Standing committee analysis. File 2007-2016 szs_ALL_ddmmyyyy.csv was not
#provided)
#15.  Checking covariates.R (This file can’t be run because it compares records used in 2017
#Standing committee analysis. File with past covariates – “covars_ddmmyyyy.csv” was not
#provided)

```


```{r echo=TRUE, message=TRUE, warning=TRUE}
# Problems encountered:
# (a) year.to and year.from commented out since set outside code
# (a) path name commented out since outside code
# (b) change year.to to year.crit.to etc
# (d) updated file name for covars etc.

source(file.path(path.Rcode,"Select final data.R"))   

```

```{r echo=TRUE, message=TRUE, warning=TRUE}
# Problems encountered:
# (a) year.to and year.from commented out since set outside code
# (a) path name commented out since outside code
# (b) change year.to to year.crit.to etc
# (d) updated file name for covars etc.
# (e) JAVA problems... have to stop

#source(file.path(path.Rcode,"sz_JAGS_model Final.R"))   # updated file name from Sz to sz

```



# code blocks yet to check out


#18.  Sz model checking.R
#19.  Raw wt_dist_jags.R
#20.  Wkd wt_dist_jags.R
#21.  Simulating raw weights.R
#22.  Simulating wkd weights.R
#23.  Covar setup_ALL_cluster.R
#24.  Trade route Calculation_All.R
#25.  Select covars cluster ALL.R
#26.  Sx x rep rate for all countries for cluster analysis Final.R
#27.  Sz_inout_tables_adj_sims.R
#28.  Mult_ctries.R
#29.  Multiple mentions all wtlimit sims.R
#30.  Creating variables for cluster analysis sims Final.R
#31.  Cluster Analysis sims Final 500.R

